---
title: "TPRS Poisson Regression with two known predictors"
author: "Samuel Herold, Kayleigh Keller"
output: html_document
---
```{r, echo=FALSE, warning=FALSE,message=FALSE}
library(spectralGP)
library(mgcv)
library(broom)
library(ggplot2)
library(dplyr)
library(readr)
```
This simulation looks at the effectiveness of varying degrees of freedom (up to ____ ) of Thin Plate Regression Splines as basis functions in a Poisson Regression to try to remove spatial confounding. The true b1 is 1, and is our target. Samples of size 1000 were taken ____ times, stored and analyzed. This time there are two known predictors that determine y. They are both added, along with the splines, to the model. 


# Trial 1
```{r, echo=FALSE,message=FALSE}
#INSERT DESIRED b1 
b1 = 1

#INSERT DATE AND TIME INTO 'dateTime'.   USE FORM:   DDMONTHYYYY-HH:MM.    Ex:   "14June2023-12:00"
#(Files found in RawData/)

dateTime <- "21June2023-15:39"


#Insert Confounder and GP settings here in form "X.XX". Include leading zero.         Ex.         "0.1"        "0.25"
#(Files found in RawData/)

G = "0.1"
C = "0.1"

estimatesRMDx = read_csv(file =paste0("../RawData/",dateTime,"-G:",G,"-C:",C,"-XEstimatesMatrixTwoPredictorsPoisson.csv"))
coverageRMDx = read_csv(file = paste0("../RawData/",dateTime,"-G:",G,"-C:",C,"-XCoverageMatrixTwoPredictorsPoisson.csv"))
standardErrorRMDx = read_csv(file = paste0("../RawData/",dateTime,"-G:",G,"-C:",C,"-XStandardErrorMatrixTwoPredictorsPoisson.csv"))
estimatesRMDx2 = read_csv(file =paste0("../RawData/",dateTime,"-G:",G,"-C:",C,"-X2EstimatesMatrixTwoPredictorsPoisson.csv"))
coverageRMDx2 = read_csv(file = paste0("../RawData/",dateTime,"-G:",G,"-C:",C,"-X2CoverageMatrixTwoPredictorsPoisson.csv"))
standardErrorRMDx2 = read_csv(file = paste0("../RawData/",dateTime,"-G:",G,"-C:",C,"-X2StandardErrorMatrixTwoPredictorsPoisson.csv"))
```

```{r, echo=FALSE}
  #MSE CALC
RMSERMDx = apply(estimatesRMDx,2,function(w) sqrt(mean((w-b1)^2)))
RMSERMDx2 = apply(estimatesRMDx2,2,function(w) sqrt(mean((w-b1)^2)))

  #Means of estimates
columnMeansRMDx = colMeans(estimatesRMDx)
columnMeansRMDx2 = colMeans(estimatesRMDx2)

  #Standard Deviation of the distribution of Estimates
sdOfEstimatesRMDx = apply(estimatesRMDx, 2, sd)
sdOfEstimatesRMDx2 = apply(estimatesRMDx2, 2, sd)

  #bias
biasRMDx = apply(estimatesRMDx, 2,function(w) mean(w-b1))
biasRMDx2 = apply(estimatesRMDx2, 2,function(w) mean(w-b1))

  #percent Coverage
percentCoverageRMDx = apply(coverageRMDx,2,mean)
percentCoverageRMDx2 = apply(coverageRMDx2,2,mean)
```

```{r, echo=FALSE}
print(paste0("X created using a spectral density of: ", G, ". Confounder created using a spectral density of: ",C,"."))


M = ncol(estimatesRMDx)+2


#Point Estimate
    ggplot() + geom_point(aes(x = 3:M,y =  columnMeansRMDx),size = .5) + geom_path(aes(3:M, columnMeansRMDx),col = "green")+ geom_point(aes(x = 3:M,y =  columnMeansRMDx2),size = .5) + geom_path(aes(3:M, columnMeansRMDx2),col = "blue")+
      labs( x = "TPRS degrees of freedom", y = "Point estimate of b1", title = "Point Estimate's of b1: Green = x1, Blue = x2",subtitle =  "by TPRS degrees of freedom") + 
      geom_hline(yintercept = b1)

#RMSE
    ggplot() +
      geom_point(aes(x = 3:M,y = RMSERMDx),size = .75) + geom_path(aes(3:M, RMSERMDx),col = "green") +
       geom_point(aes(x = 3:M,y = RMSERMDx2),size = .75) + geom_path(aes(3:M, RMSERMDx2),col = "blue") +
      labs( x = "TPRS degrees of freedom", y = "Rooot Mean Squared Error",title = "Root Mean Square Error of Point Estimates: Green = x1, Blue = x2",subtitle =  "by TPRS degrees of freedom")+ 
      geom_hline(yintercept = 0)
    

#Bias
    ggplot() +
      geom_point(aes(x = 3:M,y = biasRMDx),size = .75) + geom_path(aes(3:M, biasRMDx),col = "green") +
      geom_point(aes(x = 3:M,y = biasRMDx2),size = .75) + geom_path(aes(3:M, biasRMDx2),col = "blue") +
       labs( x = "TPRS degrees of freedom", y = "Bias",title = "Bias of Point Estimates: Green = x1, Blue = x2",subtitle =  "by TPRS degrees of freedom") + 
      geom_hline(yintercept = 0)
    

    
     
    
#Coverage
    ggplot() +
      geom_point(aes(x = 3:M,y = percentCoverageRMDx),size = .75) + geom_path(aes(3:M, percentCoverageRMDx),col = "green") +
      geom_point(aes(x = 3:M,y = percentCoverageRMDx2),size = .75) + geom_path(aes(3:M, percentCoverageRMDx2),col = "blue") +
      labs( x = "TPRS degrees of freedom", y = "Times the True b1 was Covered by Confidence Interval", title = "Coverage: Green = x1, Blue = x2",subtitle =  "by TPRS degrees of freedom")+ 
      geom_hline(yintercept = 0)

  
#SE and SD 
    ggplot() + geom_point(aes(x = 3:M,y =  colMeans(standardErrorRMDx)),size = .5) + geom_path(aes(3:M, colMeans(standardErrorRMDx)))+
      geom_point(aes(x = 3:M,y =  sdOfEstimatesRMDx),size = .5,col = "red") + geom_path(aes(3:M, sdOfEstimatesRMDx),col = "red") +
      labs( x = "TPRS degrees of freedom", y = "SD:Red and SE:Black", title = "Standard Deviation and Standard Error of x1",subtitle =  "by TPRS degrees of freedom") + 
      geom_hline(yintercept = 0)
    
      ggplot() + geom_point(aes(x = 3:M,y =  colMeans(standardErrorRMDx2)),size = .5) + geom_path(aes(3:M, colMeans(standardErrorRMDx2)))+
      geom_point(aes(x = 3:M,y =  sdOfEstimatesRMDx2),size = .5,col = "red") + geom_path(aes(3:M, sdOfEstimatesRMDx2),col = "red") +
      labs( x = "TPRS degrees of freedom", y = "SD:Red and SE:Black", title = "Standard Deviation and Standard Error of x2",subtitle =  "by TPRS degrees of freedom") + 
      geom_hline(yintercept = 0)
    


```
